EGGs are adhesive!

### 2025-03-03 calco/main.tex
minima titolo etc.

\title{On the adhesivity of EGGs}

%\titlerunning{????} %TODO optional, please use if title is longer than one line

%% Author with single affiliation.
%\author{Roberto Biondo \inst{1}{\small\Envelope}\and Davide Castelnovo\inst{1}\orcidID{0000-0002-5926-5615}
	%\and \\Fabio Gadducci\inst{1}\orcidID{0000-0003-0690-3051}
%}
%\institute{Dept.~of Computer Science, University of Pisa, Italy. 
	%\email{r.biondo@studenti.unipi.it},
	%\email{castelnovod@gmail.com},
	% \email{fabio.gadducci@unipi.it}
%}

@@

@@

\tikzset{->-/.style={decoration={
			markings,
			mark=at position #1 with {\arrow{>}}},postaction={decorate}}}

@@

In the original definition, e-graphs are defined as term graphs with an additional notion of equivalence on nodes. 
Adopting the more modern presentation via string diagrams, they are tree with sharing of subtrees, and with an 
an equivalence $\equiv$ on nodes that is closed under xxx. In plain words, if $a$ and $b$ are two constants such that
$a \equiv b$, then $f(a) \equiv f(b)$ for any unary operator $f$.

@@

\emph{Synopsis}
The paper has the following structure. 
In Section~\ref{sec:ade} we briefly recall 
the theory of
\emph{$\mathcal{M}$-adhesive categories}
and of \emph{kernel pairs}.
In Section~\ref{sec:hyper} we present the graphical structures of our interest, 
 (labelled, acyclic) hypergraphs and term graphs, and we provide a
functorial characterisation, which allows for proving their adhesivity properties.
This is expended in Section~\ref{hypereq} for describing hypergraphs 
and term graphs with equivalence and in Section~\ref{eggs} for capturing
their variants where the equivalences are closed with respect to operator application,
thus subsuming EGGs.
%
All the proofs have been moved to the appendices. 
%For the sake of completeness, 
%and in order to fix the notation, we prove all the results 
%recalled in the background section, besides those that are original 
%to our work.

>>>>

\tikzset{->-/.style={decoration={
			markings,
			mark=at position #1 with {\arrow{>}}},
			postaction={decorate}
			}}

\title{EGGs are adhesive!}
\titlerunning{EGGs are adhesive!} %TODO optional, please use if title is longer than one line

% Author with single affiliation.
\author{Roberto Biondo}
{Department of Computer Science, University of Pisa, Pisa, IT}
{r.biondo@studenti.unipi.it}{}{}

\author{Davide Castelnovo}
{Department of Computer Science, University of Pisa, Pisa, IT}
{castelnovod@gmail.com}
{https://orcid.org/0000-0002-5926-5615}{}

\author{Fabio Gadducci}
{Department of Computer Science, University of Pisa, Pisa, IT}
{fabio.gadducci@unipi.it}
{https://orcid.org/0000-0003-0690-3051}{}

@@

\Copyright{Roberto Biondo and Davide Castelnovo and Fabio Gadducci} 
\ccsdesc[500]{Theory of Computation~Models of computation}
\ccsdesc[500]{Theory of Computation~Semantics and reasoning}

\keywords{Hypergraphs, terms graphs, e-graphs, adhesive categories.} %TODO mandatory; please add comma-separated list of keywords

\funding{The research has been partially supported by the Italian MUR - Call PRIN 2022 - Project 20228KXFN2 ``Spatio-Temporal Enhancement of Neural nets for Deeply
Hierarchical Automatised Logic'' (STENDHAL) 
and by the University of Pisa - Call PRA 2022 - Project 2022\_99 ``Formal Methods for the Healthcare Domain based on Spatial Information'' (FM4HD).}

@@

@@

E-graphs can be concisely defined as (acyclic) term graphs with an additional notion of equivalence on nodes
that is closed under the operators of the signature~\cite[Section~4.2]{DetlefsNS05}.
In the presentation of term graphs via string diagrams~\cite{CastelnovoGM24}, e-graphs are trees equipped with 
the possibility of
sharing subtrees, and with an additional equivalence relation $\equiv$ on nodes that is closed under composition. 
In plain words and with a toy example:
if $a$ and $b$ are two constants such that $a \equiv b$, then $f(a) \equiv f(b)$ for any unary operator $f$.

@@

\emph{Synopsis}
The paper has the following structure. 
In Section~\ref{sec:ade} we briefly recall 
the theory of $\mathcal{M}$-adhesive categories
and of kernel pairs.
In Section~\ref{sec:hyper} we present the graphical structures of our interest, 
 (labelled, acyclic) hypergraphs and term graphs, and we provide a
functorial characterisation, which allows for proving their adhesivity properties.
This is expended in Section~\ref{hypereq} for describing hypergraphs 
and term graphs with equivalence and in Section~\ref{eggs} for capturing
their variants where the equivalences are closed with respect to operator application,
thus subsuming EGGs.
%
All the proofs have been moved to the appendices. 
%For the sake of completeness, 
%and in order to fix the notation, we prove all the results 
%recalled in the background section, besides those that are original 
%to our work.
